\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-21 metatheory-condensed.tex
Reply the comment

$$
\Rule[]
{}
{\goodWSig{\Gamma}{w\cdot}{0}}
\quad
\Rule[name=WSig-Add]
{\goodWSig{\Gamma}{\tau}{n}
  \quad \goodType{\Gamma}{A}{i}
  \quad \goodType{\Gamma, A}{B}{i}}
{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
\quad
\Rule[name=Ind-Univ]
{\goodWSig{\Gamma}{\tau}{n}}
{\goodType{\Gamma}{\bW \tau}{i}
}
$$
$$
\Rule[name=Ind-Term]
{\goodTerm{\Gamma}{T}{\bW \tau}
  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
$$
$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[\pi_1\circ\pi_1]))\; R[\pi_1\circ\pi_1]}{}}
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\tau}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
\quad j < N
}
{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
$$
$$
\Rule[name=Rec-Constr]
{ 
  % \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{T}{\bW \tau}
% \quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
\YZ{I suppose Rec-Constr is only for exhaustivity checking? If so, then would it be more intuitive to not include the w in the term? And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}\EDJreply{ 1. Yes. But not include w will cause $"Wrec"~r$ be a function type. we are not supposed to add a new intro-rule for function type(or for any other type), this is bad meta-theory design. \\
  2. It is actually the other way around -- if things are abstracted, then no beta rule can apply. More concretely, $T : \bW\tau$ is abstracted into $T' : \cU$, then this later one we cannot use $"Wrec"$ on it for sure. But once we have $\{T : \bW\tau, "rec" : \cL(RecSig~\tau)\}$ we are okay, and this, after wrap "rec" with an lambda abstraction, we will have $\cC\{T : \bW\tau, "rec" : \cL(RecSig~\tau~R), "recf" : T -> R\}$ as compilation type. Then with sealing abstraction, we will have $\Sigma(T : U, \Sigma ("rec" : ?, \Sigma("ref" : T -> R)))$ 
}\YZreply{
  What happens if the programmer forgets to seal "rec"? I suppose rec
  could still be inherited into a derived family. But it would not be
  useful if the inductive type is extended there, because rec is
  applicable only to the inductive type defined in the base family.
  Orthogonally, can you add the beta rule?
}
%\YZreply{
%  So there IS a beta rule for recursor (but you don't show it), right?
%}\EDJreply{
%  Yes. 
%  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
%  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
%  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
%}\YZreply{I'd add typeset this beta rule}\EDJreply{I just realize you use the words normalize. I will say currently this system cannot ``reduce'' any open term. But can only ``reduce'' close term. \\ 
%When the programmer program on this meta-theory, they don't need canonicity or normalization at all (they write out a chain of  judgemental equality by themselves). If they program on a proof assistant, then our proof assistant will need normalization to do type checking.
%
$$

>>>>

$$
\Rule[]
{}
{\goodWSig{\Gamma}{w\cdot}{0}}
\quad
\Rule[name=WSig-Add]
{\goodWSig{\Gamma}{\tau}{n}
  \quad \goodType{\Gamma}{A}{i}
  \quad \goodType{\Gamma, A}{B}{i}}
{\goodWSig{\Gamma}{w^+ \  \tau \  A \  B}{n+1}}
\quad
\Rule[name=Ind-Univ]
{\goodWSig{\Gamma}{\tau}{n}}
{\goodType{\Gamma}{\bW \tau}{i}
}
$$
$$
\Rule[name=Ind-Term]
{\goodTerm{\Gamma}{T}{\bW \tau}
  \quad \goodTerm{\Gamma}{a}{\pi^j_1\tau}
  \quad \goodTerm{\Gamma, \pi^j_2\tau[(id, a)]}{b}{El~("W"~T)}}
{\goodTerm{\Gamma}{"Wsup"~T~a~b}{El~("W"~T)} }
$$
$$
\Rule[name=Hdler]
{\goodType{\Gamma}{A}{}
\quad \goodType{\Gamma, A}{B}{}
\quad \goodType{\Gamma}{R}{}
}
{\goodType{\Gamma}{"RecSig1"~A~B~R \equiv \Pi A\; \Pi (\Pi B (R[\pi_1\circ\pi_1]))\; R[\pi_1\circ\pi_1]}{}}
\quad 
\Rule[name=Hdlers]
{\goodWSig{\Gamma}{\tau}{n}
\quad \goodType{\Gamma}{R}{}
}
{\goodSig{\Gamma}{"RecSig"~\tau~R}{n}}
$$
$$
\Rule[name=Hdler-Proj]
{ \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{o}{\cL("RecSig"~\tau~R)}
\quad j < N
}
{\goodTerm{\Gamma}{"prjR"~j~o}{("RecSig1"~(\pi_1^j \tau)~(\pi_2^j \tau)~R)}}
$$
$$
\Rule[name=Rec-Constr]
{ 
  % \goodWSig{\Gamma}{\tau}{N}
\quad \goodTerm{\Gamma}{T}{\bW \tau}
% \quad \goodType{\Gamma}{R}{}
\quad \goodTerm{\Gamma}{r}{\cL("RecSig"~\tau~R)}
\quad \goodTerm{\Gamma}{w}{El~("W"~T)}
}
{\goodTerm{\Gamma}{"Wrec"~r~w}{R}}
\YZ{I suppose Rec-Constr is only for exhaustivity checking? If so, then would it be more intuitive to not include the w in the term? And I suppose the programmer will want to use sealing to hide its type, or otherwise it would fail to normalize for lack of a beta rule?}\EDJreply{ 1. Yes. But not include w will cause $"Wrec"~r$ be a function type. we are not supposed to add a new intro-rule for function type(or for any other type), this is bad meta-theory design. \\
  2. It is actually the other way around -- if things are abstracted, then no beta rule can apply. More concretely, $T : \bW\tau$ is abstracted into $T' : \cU$, then this later one we cannot use $"Wrec"$ on it for sure. But once we have $\{T : \bW\tau, "rec" : \cL(RecSig~\tau)\}$ we are okay, and this, after wrap "rec" with an lambda abstraction, we will have $\cC\{T : \bW\tau, "rec" : \cL(RecSig~\tau~R), "recf" : T -> R\}$ as compilation type. Then with sealing abstraction, we will have $\Sigma(T : U, \Sigma ("rec" : ?, \Sigma("ref" : T -> R)))$ 
}\YZreply{
  What happens if the programmer forgets to seal "rec"? I suppose rec
  could still be inherited into a derived family. But it would not be
  useful if the inductive type is extended there, because rec is
  applicable only to the inductive type defined in the base family.
  Orthogonally, can you add the beta rule?
}\EDJreply{
  You are right. Since "rec" has type "\cL(RecSig~\tau~R)", after inheritance, it is basically the same type "rec : \cL(RecSig~\tau[k]~R[k])" for some "k" stands for the upcasting. but the most important $\tau[k]$ has an  unchanged shape (like the number of constructors) so it is still a recursor for inductive type of signature $\tau$. Once the parent inductive type $\bW\tau$ is extended $\bW\tau_2$ in the children family, with enriched signature $\tau_2$, then because of the signature inconsistency with $\tau[k]$, this old recursor is not useful for this new inductive type.    \\
  In other words, this recursor linkage "rec" can be used for any inductive type with signature $\tau$.\\
  Yes I will add beta rule. I see the last comment.
}
%\YZreply{
%  So there IS a beta rule for recursor (but you don't show it), right?
%}\EDJreply{
%  Yes. 
%  % https://github.com/DKXXXL/Redstone/blob/46c2420982e1fac50138924cdd7e152786cd38c8/Sketch/Extensible-Module-1/Syntax.agda#L589 \\ 
%  The only reference is here, but I commented out this url. The reduction applies to $"Wrec" \_ \_ ("Wcstr")$. \\
%  We also have canonicity (i.e. a closed term of boolean reduces to true or false), so there is definitely a beta rule for inductive type.
%}\YZreply{I'd add typeset this beta rule}\EDJreply{I just realize you use the words normalize. I will say currently this system cannot ``reduce'' any open term. But can only ``reduce'' close term. \\ 
%When the programmer program on this meta-theory, they don't need canonicity or normalization at all (they write out a chain of  judgemental equality by themselves). If they program on a proof assistant, then our proof assistant will need normalization to do type checking.
%
$$
